---
title: "Project 2 EDA"
author: 'MSDS 6372: Rob Lane'
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = T)
```
# Libraries
```{r}
# Libraries for KNN
library(class)
library(e1071)
library(caret)

# Library for ROC
library(pROC)
```
# Data
```{r}
library(aplore3)
#head(glow_bonemed)

class_colors <- c("#005d8c", "#ec9d3f")
```

glow_bonemed variables:

| **Variable** | **Type**     | **Description**                                                                                                                                     |
|--------------|--------------|-----------------------------------------------------------------------------------------------------------------------------------------------------|
| bonemed      | 2-factor     | Bone medications at enrollment (1: No, 2: Yes)                                                                                                      |
| bonemed_fu   | 2-factor     | Bone medications at follow-up (1: No, 2: Yes)                                                                                                       |
| bonetreat    | 2-factor     | Bone medications both at enrollment and follow-up (1: No, 2: Yes)                                                                                   |
| sub_id       | int          | Identification Code (1 - n)                                                                                                                         |
| site_id      | int          | Study Site (1 - 6)                                                                                                                                  |
| phy_id       | int          | Physician ID code (128 unique codes)                                                                                                                |
| priorfrac    | 2-factor     | History of Prior Fracture (1: No, 2: Yes)                                                                                                           |
| age          | int          | Age at Enrollment (Years)                                                                                                                           |
| weight       | num          | Weight at enrollment (Kilograms)                                                                                                                    |
| height       | int          | Height at enrollment (Centimeters)                                                                                                                  |
| bmi          | num          | Body Mass Index ($\frac{kg}{m^2}$)                                                                                                                  |
| premeno      | 2-factor     | Menopause before age 45 (1: No, 2: Yes)                                                                                                             |
| momfrac      | 2-factor     | Mother had hip fracture (1: No, 2: Yes)                                                                                                             |
| armassist    | 2-factor     | Arms are needed to stand from a chair (1: No, 2: Yes)                                                                                               |
| smoke        | 2-factor     | Former or current smoker (1: No, 2: Yes)                                                                                                            |
| raterisk     | 3-factor     | Self-reported risk of fracture (1: Less than others of the same age, <br>2: Same as others of the same age, 3: Greater than others of the same age) |
| fracscore    | int          | Fracture Risk Score (Composite Risk Score)                                                                                                          |
| **fracture** | **2-factor** | **Any fracture in first year (1: No, 2: Yes)**                                                                                                      |

```{r, echo=F, fig.width=7.5, fig.height=7.5}
library(graphics)
bone <- glow_bonemed

#Excluding IDs because they don't describe the problem.
exclude_vars <- c("sub_id", "site_id", "phy_id")
include_cols <- setdiff(names(bone), exclude_vars)

pairs(bone[, include_cols],
      col = class_colors[ifelse(bone$fracture=="Yes",2,1)])

```

Note: Turqoise indicates a bone fracture within the first year occured.

**Multicollinearity Found:**  
fracscore : age  
bmi : weight  
site_id : phy_id  *Because the IDs don't really tell much without factoring, and because there are a high number of factors within this class, these last two should be omitted from the model.  The code above reflects this change.*


```{r proportions_list, message=F, fig.width = 7, fig.height = 7}
library(ggplot2)
library(dplyr)
library(patchwork)


# Step 1: Create a list of factor variables
factor_list <- c("bonemed", "bonemed_fu", "bonetreat", "priorfrac", "premeno", "momfrac", "armassist", "smoke", "raterisk")

# Create an empty list to store the plots
plots_list <- list()

# Step 2: Create a loop to generate plots for each variable and store them in the list
for (variable in factor_list) {
  g <- bone %>%
    group_by(fracture, !!as.name(variable)) %>%
    summarise(cnt = n()) %>%
    mutate(perc = round(cnt / sum(cnt), 4))
  
  plot <- ggplot(g, aes(x = !!as.name(variable), y = perc, colour = fracture)) +
    geom_bar(aes(fill = fracture), stat = "identity", position = "dodge") +
    scale_colour_manual(values = class_colors) +
    scale_fill_manual(values = class_colors) +
    ylab("Proportion") +
    labs(title = paste("Variable:", variable)) +# Add a title to each plot
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +# Rotate x-axis labels by 45 degrees
    guides(fill=F, color=F) # Hide individual legends for each plot since they're all the same.
  
  # Store the plot in the list
  plots_list[[variable]] <- plot
}

# Combine the plots into a 3x3 grid using patchwork
grid_plot <- wrap_plots(plotlist = plots_list, ncol = 3)

# Create a separate plot for the legend
legend_plot <- ggplot(bone, aes(x = "", y = "", fill = fracture)) +
  geom_bar(stat = "identity", position = "dodge", width=.5, just=.1) +
  scale_fill_manual(values = class_colors) +
  guides(fill = guide_legend(title = "Fracture", keywidth = 0.7, keyheight = 0.5, size = 2)) + # Customize the legend title, size, and key size
  theme_void() + # Remove unnecessary elements from the plot, leaving only the legend
  theme(legend.text = element_text(size = 8)) # Adjust the size of the legend text

# Combine the grid of plots and the legend plot using patchwork
final_plot <- grid_plot + legend_plot

# Display the final combined plot
print(final_plot)
```

```{r loess_list, fig.width=6, fig.height=4, message=F}
  #class_colors <- c("No" = "salmon", "Yes" = "turquoise")

# Step 1: Create a list of factor variables
numeric_list <- c("age", "weight", "height", "bmi", "fracscore")

# Create an empty list to store the plots
plots_list <- list()

for (variable in numeric_list) 
{
  plot <- ggplot(bone, aes(x = !!as.name(variable), y = ifelse(fracture == "Yes",1,0))) +
            geom_point() +
            geom_smooth(method = "loess", size = 1, span = 0.75) +
            ylim(-0.2, 1.2) +
            ylab("Fracture = 1")
  plots_list[[variable]] <- plot
  

}

# plots_list[["legend"]] <- ggplot(bone, aes(x = "", y = "", fill = fracture)) +
#   geom_bar(stat = "identity", position = "dodge", width=.5, just=.1) +
#   scale_fill_manual(values = class_colors) +
#   guides(fill = guide_legend(title = "Fracture", keywidth = 0.7, keyheight = 0.5, size = 2)) + # Customize the legend title, size, and key size
#   theme_void() + # Remove unnecessary elements from the plot, leaving only the legend
#   theme(legend.text = element_text(size = 8)) # Adjust the size of the legend text

grid_plot <- wrap_plots(plotlist = plots_list, ncol = 3)

print(grid_plot)
```

```{r loess_interactions, message=F, eval=F}
color_scale <- c("red","blue","black")


i=0
plots_list <- list()

for (factor in factor_list){
  for (numeric in numeric_list) {
    i <- i +1
    plot <- ggplot(bone, aes(x = !!as.name(numeric), y = ifelse(fracture == "Yes",1,0), colour=factor(!!as.name(factor)))) +
              geom_point() +
              geom_smooth(method = "loess", size = 1, span = 1) +
              ylim(-0.2, 1.2) +
              ylab("Fracture = 1") +
              scale_colour_manual(values = color_scale) +
              labs(title=paste0("Interaction #",i))
    
    print(plot)
    #plots_list[[variable]] <- plot
    #plots_list[[paste(factor, numeric, sep = "_")]] <- plot
  }
}

# Combine the plots into a grid using patchwork
#grid_plot <- wrap_plots(plotlist = plots_list, ncol = 3)

#print(grid_plot)
```

```{r loess_interactions_2, message=F, fig.width=7, fig.height=9}
# library(ggplot2)
# library(dplyr)
# library(patchwork)

# This is not intended to be the final list of interactions, just those that seemed worthy of further exploration from above.
interactions <-c(
  # factor, numeric
  c("bonemed",    "weight"), #2
  c("bonemed",    "bmi"),    #4
  c("bonemed_fu", "weight"), #7
  c("bonemed_fu", "height"), #8
  c("bonemed_fu", "bmi"),    #9
  c("bonetreat",  "weight"),#12
  c("bonetreat",  "bmi"),   #14
  c("priorfrac",  "weight"),#17
  c("priorfrac",  "bmi"),   #19
  c("momfrac",    "age"),   #26
  c("momfrac",    "weight"),#27
  c("momfrac",    "bmi"),   #29
  c("smoke",      "height"),#38
  c("raterisk",   "weight") #42
)

inter_nums <- c(2, 4, 7, 8, 9, 12, 14, 17, 19, 26, 27, 29, 38, 42)

color_scale <- c("red", "blue", "black")

i <- 0
plots_list <- list()
for (factor in factor_list) {
  for (numeric in numeric_list) {
    i <- i + 1
    if (i %in% inter_nums) {
      # Remove NAs before creating the plot
      bone_plot <- na.omit(bone)
      
      plot <- ggplot(bone_plot, aes(x = !!as.name(numeric), y = ifelse(fracture == "Yes", 1, 0), colour = factor(!!as.name(factor)))) +
        geom_point() +
        geom_smooth(method = "loess", size = 1, span = 1) +
        ylim(-0.2, 1.2) +
        ylab("Fracture = 1") + 
        xlab(paste(numeric," vs. ", factor,".fac",sep="")) +
        scale_colour_manual(values = color_scale) +
        guides(fill=F, color=F) # Hide individual legends for each plot 

      
      plots_list[[paste(factor, numeric, sep = "_")]] <- plot
    }
  }
}

# Combine the plots into a grid using patchwork
grid_plot <- wrap_plots(plotlist = plots_list, ncol = 3)

print(grid_plot)


#Note: when momfrac has an interaction, the numeric may have a poly(2) term
```
# KNN
```{r KNN prediction models}
# Set factors as binaries
bone$priorfrac_num<-ifelse(bone$priorfrac=="Yes",1,0)
bone$premeno_num<-ifelse(bone$premeno=="Yes",1,0)
bone$momfrac_num<-ifelse(bone$momfrac=="Yes",1,0)
bone$armassist_num<-ifelse(bone$armassist=="Yes",1,0)
bone$smoke_num<-ifelse(bone$smoke=="Yes",1,0)
bone$raterisk_num<-ifelse(bone$raterisk=="Less",0,ifelse(bone$raterisk=="Same",1,2))
bone$bonemed_num<-ifelse(bone$bonemed=="Yes",1,0)
bone$bonemed_fu_num<-ifelse(bone$bonemed_fu=="Yes",1,0)
bone$bonetreat_num<-ifelse(bone$bonetreat=="Yes",1,0)
bone$frac_num<-ifelse(bone$fracture=="Yes",1,0)

# Split our data into a train and test set
set.seed(123)
splitPerc = .7
trainIndices = sample(1:dim(bone)[1],round(splitPerc * dim(bone)[1]))
dftrain = bone[trainIndices,]
dftest = bone[-trainIndices,]

# k = 3 for 3 Continuous Variables with Strong Evidence via Ttest
fitControl<-trainControl(method="repeatedcv",number=10,repeats=1,classProbs=TRUE, summaryFunction=mnLogLoss)
knn_model1 <-train(fracture ~ age+height+fracscore,
                           data=dftrain,
                    method="knn",
                    trControl=fitControl,
                    tuneLength=3,
                             metric="logLoss")
classifications = knn(dftrain[,c(5,7,14)],dftest[,c(5,7,14)],dftrain$fracture, 
                      prob = TRUE, k = 3)
table(classifications,dftest$fracture)
confusionMatrix(table(classifications,dftest$fracture), positive = "Yes")

# k = 15 for 3 Continuous Variables with Strong Evidence via Ttest
fitControl<-trainControl(method="repeatedcv",number=10,repeats=1,classProbs=TRUE, summaryFunction=mnLogLoss)
knn_model2 <-train(fracture ~ age+height+fracscore,
                           data=dftrain,
                    method="knn",
                    trControl=fitControl,
                    tuneLength=15,
                             metric="logLoss")
classifications = knn(dftrain[,c(5,7,14)],dftest[,c(5,7,14)],dftrain$fracture, 
                      prob = TRUE, k = 15)
table(classifications,dftest$fracture)
confusionMatrix(table(classifications,dftest$fracture), positive = "Yes")

# k = 15 for all variables
fitControl<-trainControl(method="repeatedcv",number=10,repeats=1,classProbs=TRUE, summaryFunction=mnLogLoss)
knn_model3 <-train(fracture ~ age+weight+height+bmi+
                     fracscore+priorfrac_num+premeno_num+
                     momfrac_num+armassist_num+smoke_num+bonemed_num+
                     bonemed_fu_num+bonetreat_num+raterisk_num,
                           data=dftrain,
                    method="knn",
                    trControl=fitControl,
                    tuneLength=15,
                             metric="logLoss")
classifications = knn(dftrain[,c(5:8,14,19:27)],dftest[,c(5:8,14,19:27)],dftrain$fracture, 
                      prob = TRUE, k = 15)
table(classifications,dftest$fracture)
confusionMatrix(table(classifications,dftest$fracture), positive = "Yes")

# k = 15 for Custom Complex Model
fitControl<-trainControl(method="repeatedcv",number=10,repeats=1,classProbs=TRUE, summaryFunction=mnLogLoss)
knn_model4 <-train(fracture ~ age+height+priorfrac_num+
                     momfrac_num+armassist_num+bonemed_num+
                     bonemed_fu_num+bonetreat_num+raterisk_num,
                           data=dftrain,
                    method="knn",
                    trControl=fitControl,
                    tuneLength=15,
                             metric="logLoss")
classifications = knn(dftrain[,c(7,19,21,22,24:27)],dftest[,c(7,19,21,22,24:27)],dftrain$fracture, 
                      prob = TRUE, k = 15)
table(classifications,dftest$fracture)
confusionMatrix(table(classifications,dftest$fracture), positive = "Yes")
```


```{r Loop to find best k}
# Continuous Variables
# Loop for many k and the average of many training / test partition
iterations = 500
numks = 30

masterAcc = matrix(nrow = iterations, ncol = numks)
  
for(j in 1:iterations)
{
accs = data.frame(accuracy = numeric(30), k = numeric(30))
trainIndices = sample(1:dim(bone)[1],round(splitPerc * dim(bone)[1]))
train = bone[trainIndices,]
test = bone[-trainIndices,]
for(i in 1:numks)
{
  classifications = knn(train[,c(5,7,14)],test[,c(5,7,14)],train$fracture, prob = TRUE, k = i)
  table(classifications,test$fracture)
  CM = confusionMatrix(table(classifications,test$fracture), positive = "Yes")
  masterAcc[j,i] = CM$overall[1]
}

}

MeanAcc = colMeans(masterAcc)

plot(seq(1,numks,1),MeanAcc, type = "l")

# Continuous & Categorical Variables
# Loop for many k and the average of many training / test partition
iterations = 500
numks = 30

masterAcc = matrix(nrow = iterations, ncol = numks)
  
for(j in 1:iterations)
{
accs = data.frame(accuracy = numeric(30), k = numeric(30))
trainIndices = sample(1:dim(bone)[1],round(splitPerc * dim(bone)[1]))
train = bone[trainIndices,]
test = bone[-trainIndices,]
for(i in 1:numks)
{
  classifications = knn(train[,c(5:8,14,19:27)],test[,c(5:8,14,19:27)],train$fracture, prob = TRUE, k = i)
  table(classifications,test$fracture)
  CM = confusionMatrix(table(classifications,test$fracture), positive = "Yes")
  masterAcc[j,i] = CM$overall[1]
}

}

MeanAcc = colMeans(masterAcc)

plot(seq(1,numks,1),MeanAcc, type = "l")
```

Let's run a Cross Validation on the model:

```{r}
#Internal CV for continuous variables, K = 3
classifications = knn.cv(bone[,c(5,7,14)],bone$fracture,prob = TRUE, k = 3)
confusionMatrix(table(classifications,bone$fracture), positive = "Yes")

#Internal CV for continuous variables, K = 15
classifications = knn.cv(bone[,c(5,7,14)],bone$fracture,prob = TRUE, k = 15)
confusionMatrix(table(classifications,bone$fracture), positive = "Yes")

#Internal CV for all variables
classifications = knn.cv(bone[,c(5:8,14,19:27)],bone$fracture,prob = TRUE, k = 15)
confusionMatrix(table(classifications,bone$fracture), positive = "Yes")

#Internal CV for custom model
classifications = knn.cv(bone[,c(5,7,19,21:22,24:27)],bone$fracture,prob = TRUE, k = 15)
confusionMatrix(table(classifications,bone$fracture), positive = "Yes")

```

Predicted Probabilities for KNN Models

```{r}
# knn_model1
knn1.predprobs <- predict(knn_model1, dftrain, type = "prob")[,"Yes"]
knn1.roc <- roc(response=dftrain$fracture, predictor = knn1.predprobs, levels = c("No", "Yes"))

plot(knn1.roc, col = "#990943", print.thres="best",lwd = 2, main = "ROC Curves Comparison")
cat("KNN AUROC\n")
auc(knn1.roc)

# knn_model2
knn2.predprobs <- predict(knn_model2, dftrain, type = "prob")[,"Yes"]
knn2.roc <- roc(response=dftrain$fracture, predictor = knn2.predprobs, levels = c("No", "Yes"))

plot(knn2.roc, col = "#08A338", print.thres="best",lwd = 2, main = "ROC Curves Comparison")
cat("KNN AUROC\n")
auc(knn2.roc)

# knn_model3
knn3.predprobs <- predict(knn_model3, dftrain, type = "prob")[,"Yes"]
knn3.roc <- roc(response=dftrain$fracture, predictor = knn3.predprobs, levels = c("No", "Yes"))

plot(knn3.roc, col = "#3908A3", print.thres="best",lwd = 2, main = "ROC Curves Comparison")
cat("KNN AUROC\n")
auc(knn3.roc)

# knn_model4
knn4.predprobs <- predict(knn_model4, dftrain, type = "prob")[,"Yes"]
knn4.roc <- roc(response=dftrain$fracture, predictor = knn4.predprobs, levels = c("No", "Yes"))

plot(knn4.roc, col = "salmon", print.thres="best",lwd = 2, main = "ROC Curves Comparison")
cat("KNN AUROC\n")
auc(knn4.roc)

# All 4 Models
plot(knn1.roc, col = "#990943", print.thres="best",lwd = 2, main = "ROC Curves Comparison")
lines(knn2.roc, col = "#08A338", lwd = 2)
lines(knn3.roc, col = "#3908A3", lwd = 2)
lines(knn4.roc, col = "salmon", lwd = 2)
legend("bottomright", legend = c("K=3, Cont. Var", "K=15, Cont. Var","K=15, All variables","K=15, Custom Complex Model"), col = c("#990943", "#08A338","#3908A3","salmon"), lwd = 2)
```

Inspect Probabilities

Source: https://shihchingfu.github.io/knn-caret-example/#
```{r}
class_colors <- c("#005d8c", "#ec9d3f")

# Knn Model 1
knn1.probabilities <- dftrain %>%
  mutate(Predicted_prob = knn1.predprobs)

knn1.probabilities %>%
  ggplot() +
  aes(x = Predicted_prob, fill = fracture) +
  geom_histogram(bins = 20) +
  labs(x = "Probability", y = "Count", title = "Distribution of predicted probabilities" ) +
  scale_color_manual(values = class_colors) +
  scale_fill_manual(values = class_colors)

# Knn Model 2
knn2.probabilities <- dftrain %>%
  mutate(Predicted_prob = knn2.predprobs)

knn2.probabilities %>%
  ggplot() +
  aes(x = Predicted_prob, fill = fracture) +
  geom_histogram(bins = 20) +
  labs(x = "Probability", y = "Count", title = "Distribution of predicted probabilities" ) +
  scale_color_manual(values = class_colors) +
  scale_fill_manual(values = class_colors)

# Knn Model 3
knn3.probabilities <- dftrain %>%
  mutate(Predicted_prob = knn1.predprobs)

knn3.probabilities %>%
  ggplot() +
  aes(x = Predicted_prob, fill = fracture) +
  geom_histogram(bins = 20) +
  labs(x = "Probability", y = "Count", title = "Distribution of predicted probabilities" ) +
  scale_color_manual(values = class_colors) +
  scale_fill_manual(values = class_colors)

# Knn Model 4
knn4.probabilities <- dftrain %>%
  mutate(Predicted_prob = knn1.predprobs)

knn4.probabilities %>%
  ggplot() +
  aes(x = Predicted_prob, fill = fracture) +
  geom_histogram(bins = 20) +
  labs(x = "Probability", y = "Count", title = "Distribution of predicted probabilities" ) +
  scale_color_manual(values = class_colors) +
  scale_fill_manual(values = class_colors)
```

Fine tuning KNN threshholds

Source: https://shihchingfu.github.io/knn-caret-example/#
```{r}
# KNN Model 1
knn1.probabilities <- knn1.probabilities %>%
  mutate(class = ifelse(Predicted_prob > 0.24, "Yes", "No")) %>%
  mutate(class = factor(class))
confusionMatrix(table(knn1.probabilities$class,dftrain$fracture), positive = "Yes")

# KNN Model 2
knn2.probabilities <- knn2.probabilities %>%
  mutate(class = ifelse(Predicted_prob > 0.282, "Yes", "No")) %>%
  mutate(class = factor(class))
confusionMatrix(table(knn2.probabilities$class,dftrain$fracture), positive = "Yes")

# KNN Model 3
knn3.probabilities <- knn3.probabilities %>%
  mutate(class = ifelse(Predicted_prob > 0.209, "Yes", "No")) %>%
  mutate(class = factor(class))
confusionMatrix(table(knn3.probabilities$class,dftrain$fracture), positive = "Yes")

# KNN Model 4
knn4.probabilities <- knn4.probabilities %>%
  mutate(class = ifelse(Predicted_prob > 0.262, "Yes", "No")) %>%
  mutate(class = factor(class))
confusionMatrix(table(knn4.probabilities$class,dftrain$fracture), positive = "Yes")



```
# Objective 1

```{r}
library(jtools)
library(sjPlot)
library(ResourceSelection)

# Simple Logistic Regression

# Significant Continuous Variables only:
logistic1<-glm(fracture~age + height + fracscore,data=bone, family="binomial")
summary(logistic1)

# Transform for interpretation
exp(coef(logistic1))
exp(confint(logistic1))

#jtools
effect_plot(logistic1, pred = age, interval = TRUE, plot.points = TRUE, 
            jitter = 0.05)

#sjPlot
plot_model(logistic1,type="pred",terms=c("age","height","fracscore"))
```
Feature Selection
```{r}
# Feature Selection for Simple Model (No interactions or polynomials)
fitControl<-trainControl(method="repeatedcv",number=5,repeats=1,classProbs=TRUE, summaryFunction=mnLogLoss)
set.seed(1234)
#note CV and error metric are not really used here, but logLoss is reported for the final model.
Step.fit<-train(fracture~age+weight+height+bmi+priorfrac+premeno+momfrac+armassist+
                  smoke+raterisk+fracscore+bonemed+bonemed_fu+bonetreat,
                    data=bone,
                    method="glmStepAIC",
                    trControl=fitControl,
                    metric="logLoss")
coef(Step.fit$finalModel)
```
Best AIC Model:
```{r Custom_Models}

# Using variables from best AIC at 405.57
logistic2<-glm(fracture~ weight + bmi + momfrac + raterisk + fracscore + 
                 bonemed + bonemed_fu + bonetreat,data=dftrain, family="binomial")
summary(logistic2)

plot_model(logistic2,type="pred",terms=c("fracscore","weight","bmi","raterisk","momfrac","bonemed","bonemed_fu","bonetreat"))

# Transform for interpretation
exp(coef(logistic2))
exp(confint(logistic2))

# Removing insignificant variables
logistic3<-glm(fracture~ weight + bmi + fracscore + 
                 bonemed + bonemed_fu + bonetreat,data=bone, family="binomial")
summary(logistic3)

# Transform for interpretation
exp(coef(logistic3))
exp(confint(logistic3))
```
Probabilities and Predictions
```{r}
class2_colors <- c("#990943", "#08A338","#3908A3","salmon")

# Simple Logistic Regression, Significant Continuous Variables
logistic1.predprobs <- predict(logistic1, bone, type = "response")
logistic1.roc <- roc(response=bone$fracture, predictor = logistic1.predprobs, levels = c("No", "Yes"))

#plot(logistic1.roc, col = class2_colors[1], print.thres="best",lwd = 2, main = "ROC Curves Comparison")
cat("Age, Height & Frac Score AUROC\n")
auc(logistic1.roc)

# Custom Logistic Regression after Feature Selection
logistic2.predprobs <- predict(logistic2, bone, type = "response")
logistic2.roc <- roc(response=bone$fracture, predictor = logistic2.predprobs, levels = c("No", "Yes"))

#plot(logistic2.roc, col = class2_colors[2], print.thres="best",lwd = 2, main = "ROC Curves Comparison")
cat("Custom Model 2 AUROC\n")
auc(logistic2.roc)

# Second Custom Logistic Regression after Feature Selection 
logistic3.predprobs <- predict(logistic3, bone, type = "response")
logistic3.roc <- roc(response=bone$fracture, predictor = logistic3.predprobs, levels = c("No", "Yes"))

#plot(logistic3.roc, col = class2_colors[3], print.thres="best",lwd = 2, main = "ROC Curves Comparison")
cat("Custom Model 3 AUROC\n")
auc(logistic3.roc)

# All 3 Models
plot(logistic1.roc, col = class2_colors[1], print.thres="best",lwd = 2, main = "ROC Curves Comparison")
lines(logistic2.roc, col = class2_colors[2], lwd = 2)
lines(logistic3.roc, col = class2_colors[3], lwd = 2)
legend("bottomright", legend = c("Age, Height, & Frac Score", "Custom Model 1","Custom Model 2"), col = class2_colors, lwd = 2)
```
Confusion Matrices for Simple Logistic Regression
```{r}
# Simple Logistic Regression, Significant Continuous Variables
logistic1.predprobs <- data.frame(as.numeric(logistic1.predprobs))

logistic1.predprobs <- logistic1.predprobs %>%
  mutate(class = ifelse(logistic1.predprobs[,1] > 0.284, "Yes", "No")) %>%
  mutate(class = factor(class))
confit1 <- confusionMatrix(table(logistic1.predprobs$class,bone$fracture), positive = "Yes")

# Custom Logistic Regression after Feature Selection
logistic2.predprobs <- data.frame(as.numeric(logistic2.predprobs))

logistic2.predprobs <- logistic2.predprobs %>%
  mutate(class = ifelse(logistic1.predprobs[,1] > 0.221, "Yes", "No")) %>%
  mutate(class = factor(class))
confit2 <- confusionMatrix(table(logistic2.predprobs$class,bone$fracture), positive = "Yes")

# Second Custom Logistic Regression after Feature Selection
logistic3.predprobs <- data.frame(as.numeric(logistic3.predprobs))

logistic3.predprobs <- logistic3.predprobs %>%
  mutate(class = ifelse(logistic3.predprobs[,1] > 0.281, "Yes", "No")) %>%
  mutate(class = factor(class))
confit3 <- confusionMatrix(table(logistic3.predprobs$class,bone$fracture), positive = "Yes")

cat("Confusion Matrices for Simple Logistic Regression Models\n")
cat("### Age, Height & Frac Score ###          ##### Custom Model 1 ######          ##### Custom Model 2 ######")
confit1
confit2
confit3
```










