---
title: "Project 2 EDA"
author: 'MSDS 6372: Rob Lane'
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

TESTING IF THIS PUTS IT IN MY BRANCH
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

```{r}
library(aplore3)
#head(glow_bonemed)
```

glow_bonemed variables:

| **Variable** | **Type**     | **Description**                                                                                                                                     |
|--------------|--------------|-----------------------------------------------------------------------------------------------------------------------------------------------------|
| bonemed      | 2-factor     | Bone medications at enrollment (1: No, 2: Yes)                                                                                                      |
| bonemed_fu   | 2-factor     | Bone medications at follow-up (1: No, 2: Yes)                                                                                                       |
| bonetreat    | 2-factor     | Bone medications both at enrollment and follow-up (1: No, 2: Yes)                                                                                   |
| sub_id       | int          | Identification Code (1 - n)                                                                                                                         |
| site_id      | int          | Study Site (1 - 6)                                                                                                                                  |
| phy_id       | int          | Physician ID code (128 unique codes)                                                                                                                |
| priorfrac    | 2-factor     | History of Prior Fracture (1: No, 2: Yes)                                                                                                           |
| age          | int          | Age at Enrollment (Years)                                                                                                                           |
| weight       | num          | Weight at enrollment (Kilograms)                                                                                                                    |
| height       | int          | Height at enrollment (Centimeters)                                                                                                                  |
| bmi          | num          | Body Mass Index ($\frac{kg}{m^2}$)                                                                                                                  |
| premeno      | 2-factor     | Menopause before age 45 (1: No, 2: Yes)                                                                                                             |
| momfrac      | 2-factor     | Mother had hip fracture (1: No, 2: Yes)                                                                                                             |
| armassist    | 2-factor     | Arms are needed to stand from a chair (1: No, 2: Yes)                                                                                               |
| smoke        | 2-factor     | Former or current smoker (1: No, 2: Yes)                                                                                                            |
| raterisk     | 3-factor     | Self-reported risk of fracture (1: Less than others of the same age, <br>2: Same as others of the same age, 3: Greater than others of the same age) |
| fracscore    | int          | Fracture Risk Score (Composite Risk Score)                                                                                                          |
| **fracture** | **2-factor** | **Any fracture in first year (1: No, 2: Yes)**                                                                                                      |

```{r, echo=F, fig.width=7.5, fig.height=7.5}
library(graphics)
bone <- glow_bonemed
#Survived = 1, Perished = 0.  Therefore, #005d8c means perished, #ec9d3f means survived.
class_colors <- c("#005d8c", "#ec9d3f")

#Excluding IDs because they don't describe the problem.
exclude_vars <- c("sub_id", "site_id", "phy_id")
include_cols <- setdiff(names(bone), exclude_vars)

pairs(bone[, include_cols],
      col = class_colors[ifelse(bone$fracture=="Yes",2,1)])

```

Note: Turqoise indicates a bone fracture within the first year occured.

**Multicollinearity Found:**  
fracscore : age  
bmi : weight  
site_id : phy_id  *Because the IDs don't really tell much without factoring, and because there are a high number of factors within this class, these last two should be omitted from the model.  The code above reflects this change.*


```{r proportions_list, message=F, fig.width = 7, fig.height = 7}
library(ggplot2)
library(dplyr)
library(patchwork)


# Step 1: Create a list of factor variables
factor_list <- c("bonemed", "bonemed_fu", "bonetreat", "priorfrac", "premeno", "momfrac", "armassist", "smoke", "raterisk")

# Create an empty list to store the plots
plots_list <- list()

# Step 2: Create a loop to generate plots for each variable and store them in the list
for (variable in factor_list) {
  g <- bone %>%
    group_by(fracture, !!as.name(variable)) %>%
    summarise(cnt = n()) %>%
    mutate(perc = round(cnt / sum(cnt), 4))
  
  class_colors <- c("No" = "#005d8c", "Yes" = "#ec9d3f")
  
  plot <- ggplot(g, aes(x = !!as.name(variable), y = perc, colour = fracture)) +
    geom_bar(aes(fill = fracture), stat = "identity", position = "dodge") +
    scale_colour_manual(values = class_colors) +
    scale_fill_manual(values = class_colors) +
    ylab("Proportion") +
    labs(title = paste("Variable:", variable)) +# Add a title to each plot
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +# Rotate x-axis labels by 45 degrees
    guides(fill=F, color=F) # Hide individual legends for each plot since they're all the same.
  
  # Store the plot in the list
  plots_list[[variable]] <- plot
}

# Combine the plots into a 3x3 grid using patchwork
grid_plot <- wrap_plots(plotlist = plots_list, ncol = 3)

# Create a separate plot for the legend
legend_plot <- ggplot(bone, aes(x = "", y = "", fill = fracture)) +
  geom_bar(stat = "identity", position = "dodge", width=.5, just=.1) +
  scale_fill_manual(values = class_colors) +
  guides(fill = guide_legend(title = "Fracture", keywidth = 0.7, keyheight = 0.5, size = 2)) + # Customize the legend title, size, and key size
  theme_void() + # Remove unnecessary elements from the plot, leaving only the legend
  theme(legend.text = element_text(size = 8)) # Adjust the size of the legend text

# Combine the grid of plots and the legend plot using patchwork
final_plot <- grid_plot + legend_plot

# Display the final combined plot
print(final_plot)
```

```{r loess_list, fig.width=6, fig.height=4, message=F}
  #class_colors <- c("No" = "#005d8c", "Yes" = "#ec9d3f")
library(ggplot2)
library(dplyr)
library(patchwork)

# Step 1: Create a list of factor variables
numeric_list <- c("age", "weight", "height", "bmi", "fracscore")

# Create an empty list to store the plots
plots_list <- list()

for (variable in numeric_list) 
{
  plot <- ggplot(bone, aes(x = !!as.name(variable), y = ifelse(fracture == "Yes",1,0))) +
            geom_point() +
            geom_smooth(method = "loess", size = 1, span = 0.75) +
            ylim(-0.2, 1.2) +
            ylab("Fracture = 1")
  plots_list[[variable]] <- plot
  

}

# plots_list[["legend"]] <- ggplot(bone, aes(x = "", y = "", fill = fracture)) +
#   geom_bar(stat = "identity", position = "dodge", width=.5, just=.1) +
#   scale_fill_manual(values = class_colors) +
#   guides(fill = guide_legend(title = "Fracture", keywidth = 0.7, keyheight = 0.5, size = 2)) + # Customize the legend title, size, and key size
#   theme_void() + # Remove unnecessary elements from the plot, leaving only the legend
#   theme(legend.text = element_text(size = 8)) # Adjust the size of the legend text

grid_plot <- wrap_plots(plotlist = plots_list, ncol = 3)

print(grid_plot)
```

```{r loess_interactions, message=F, eval=F}
color_scale <- c("red","blue","black")


i=0
plots_list <- list()

for (factor in factor_list){
  for (numeric in numeric_list) {
    i <- i +1
    plot <- ggplot(bone, aes(x = !!as.name(numeric), y = ifelse(fracture == "Yes",1,0), colour=factor(!!as.name(factor)))) +
              geom_point() +
              geom_smooth(method = "loess", size = 1, span = 1) +
              ylim(-0.2, 1.2) +
              ylab("Fracture = 1") +
              scale_colour_manual(values = color_scale) +
              labs(title=paste0("Interaction #",i))
    
    print(plot)
    #plots_list[[variable]] <- plot
    #plots_list[[paste(factor, numeric, sep = "_")]] <- plot
  }
}

# Combine the plots into a grid using patchwork
#grid_plot <- wrap_plots(plotlist = plots_list, ncol = 3)

#print(grid_plot)
```

```{r loess_interactions_2, message=F, fig.width=7, fig.height=9}
# library(ggplot2)
# library(dplyr)
# library(patchwork)

# This is not intended to be the final list of interactions, just those that seemed worthy of further exploration from above.
interactions <-c(
  # factor, numeric
  c("bonemed",    "weight"), #2
  c("bonemed",    "bmi"),    #4
  c("bonemed_fu", "weight"), #7
  c("bonemed_fu", "height"), #8
  c("bonemed_fu", "bmi"),    #9
  c("bonetreat",  "weight"),#12
  c("bonetreat",  "bmi"),   #14
  c("priorfrac",  "age"),   #16
  c("priorfrac",  "weight"),#17
  c("priorfrac",  "bmi"),   #19
  c("momfrac",    "age"),   #26
  c("momfrac",    "weight"),#27
  c("momfrac",    "bmi"),   #29
  c("smoke",      "height"),#38
  c("raterisk",   "weight") #42
)

inter_nums <- c(2, 4, 7, 8, 9, 12, 14, 16, 17, 19, 26, 27, 29, 38, 42)

color_scale <- c("red", "blue", "black")

i <- 0
plots_list <- list()
for (factor in factor_list) {
  for (numeric in numeric_list) {
    i <- i + 1
    if (i %in% inter_nums) {
      # Remove NAs before creating the plot
      bone_plot <- na.omit(bone)
      
      plot <- ggplot(bone_plot, aes(x = !!as.name(numeric), y = ifelse(fracture == "Yes", 1, 0), colour = factor(!!as.name(factor)))) +
        geom_point() +
        geom_smooth(method = "loess", size = 1, span = 1) +
        ylim(-0.2, 1.2) +
        ylab("Fracture = 1") + 
        xlab(paste(numeric," vs. ", factor,".fac",sep="")) +
        scale_colour_manual(values = color_scale) +
        guides(fill=F, color=F) # Hide individual legends for each plot 

      
      plots_list[[paste(factor, numeric, sep = "_")]] <- plot
    }
  }
}

# Combine the plots into a grid using patchwork
grid_plot <- wrap_plots(plotlist = plots_list, ncol = 3)

print(grid_plot)


#Note: when momfrac has an interaction, the numeric may have a poly(2) term
```

```{r test_train_split}

library(caret)
set.seed(1234)
trainIndex<-createDataPartition(bone$fracture,p=.7,list=F)  #p: proportion of data in train

training<-bone[trainIndex,]
validate<-bone[-trainIndex,]

```





# Complex Model

```{r}

mymodel<-glm(fracture~bonemed+bonemed_fu+bonetreat+priorfrac+age+weight+height
             +bmi+premeno+momfrac+armassist+smoke+raterisk+fracscore
             +bonemed:weight
             +bonemed:bmi
             +bonemed_fu:weight
             +bonemed_fu:height
             +bonemed_fu:bmi
             +bonetreat:weight
             +bonetreat:bmi
             +priorfrac:age
             +priorfrac:weight
             +priorfrac:bmi
             # +momfrac:poly(age,2)
             # +momfrac:poly(weight,2)
             # +momfrac:poly(bmi,2)
             +momfrac:age
             +momfrac:weight
             +momfrac:bmi
             +bonetreat:weight:height**2
             +smoke:height
             +raterisk:weight
             +bmi:weight:height,
             data=training,family="binomial")
summary(mymodel)

cat("\n##### e transform odds ratio #####\n")
exp(coef(mymodel))
# cat("\n##### CI @ 90% #####\n")
# exp(confint(mymodel, level=0.95))

```

```{r}

fitControl<-trainControl(method="repeatedcv",number=10,repeats=1,classProbs=TRUE, summaryFunction=mnLogLoss)
set.seed(1234)
glmnet.fit<-train(fracture~bonemed+bonemed_fu+bonetreat+priorfrac+age+weight+height
             +bmi+premeno+momfrac+armassist+smoke+raterisk+fracscore
             +bonemed:weight
             +bonemed:bmi
             +bonemed_fu:weight
             +bonemed_fu:height
             +bonemed_fu:bmi
             +bonetreat:weight
             #+bonetreat:bmi
             +bonetreat:weight:poly(height,2)
             +priorfrac:age
             +priorfrac:weight
             +priorfrac:bmi
             +momfrac:poly(age,2)
             +momfrac:poly(weight,2)
             +momfrac:poly(bmi,2)
             # +momfrac:age
             # +momfrac:weight
             # +momfrac:bmi
             +smoke:height
             +raterisk:weight
             +bmi:weight:height,
                    data=training,
                    method="glmnet",
                    trControl=fitControl,
                    metric="logLoss")
#coef(glmnet.fit$finalModel)
# This makes a huge table that is hard to read.



opt.pen<-glmnet.fit$finalModel$lambdaOpt #penalty term
  coef(glmnet.fit$finalModel,opt.pen)
  
```

```{r}

fitControl<-trainControl(method="repeatedcv",number=5,repeats=1,classProbs=TRUE, summaryFunction=mnLogLoss)


complex_model<-train(fracture~
               
               height
             +bonemed     # Difference in ratios
             +bonemed_fu  # Difference in ratios
             +bonetreat   # Difference in ratios
             +priorfrac   # Difference in ratios
             # +age
             # +weight
             # +bmi
             # +premeno
             +momfrac
             +armassist   # Difference in ratios
             # +smoke
             +raterisk    # Difference in ratios
             # +fracscore
             # +bonemed:weight
             # +bonemed:bmi
             # +bonemed_fu:weight
             # +bonemed_fu:height
             +bonemed_fu:bmi
             # +bonetreat:weight
             # +bonetreat:bmi
             # +priorfrac:age
             # +priorfrac:weight
             +priorfrac:bmi
             # +momfrac:poly(age,2)
             +momfrac:poly(weight,2)
             # +momfrac:poly(bmi,2)
             # +momfrac:age
             # +momfrac:weight
             # +momfrac:bmi
             # +smoke:height
             +raterisk:weight
             # +bmi:weight:height
             ,
             data=training,
             method="glmnet",
             trControl=fitControl,
             metric="logLoss")
#summary(complex_model)

# cat("\n##### e transform odds ratio #####\n")
# exp(coef(complex_model))
# cat("\n##### CI @ 90% #####\n")
# exp(confint(complex_model, level=0.95))

```

```{r}
library(pROC)
complex_predictions <- predict(complex_model, training, type = "prob")[,"Yes"]

# use ROC to help tune the threshold
complex.roc <- roc(response = training$fracture, predictor = complex_predictions,
                  levels = c("No", "Yes"))

# Compute the threshold for maximum sensitivity
max_sens_threshold <- complex.roc$thresholds[which.max(complex.roc$sensitivities)]
cat("Max Sensitivity Threshold: ", max_sens_threshold,"\n")
max_spec_threshold <- complex.roc$thresholds[which.max(complex.roc$specificities)]
cat("Max Specificity Threshold: ", max_spec_threshold,"\n")
best_threshold <- coords(complex.roc, "best")$threshold
cat("Best Overall Threshold:    ", best_threshold,"\n\n\n")

# Plot the ROC curve and add a vertical line at max sensitivity threshold
plot(complex.roc, col = "blue",print.thres="best", lwd = 2, main = "ROC Curve with Max Sensitivity Threshold")


#abline(v = best_threshold, col = "red", lty = 2)
legend("bottomright", legend = c("Complex Model", "Max Sensitivity Threshold"), col = c("blue", "red"), lwd = 2)



cat("GLMNET AUROC\n")
auc(complex.roc)
#cat("KNN AUROC\n")
#auc(knn.roc)

#Getting confusion matrix
threshold = .3
complex.preds <- factor(ifelse(predictions > threshold, "No", "Yes"),
                    levels = c("No", "Yes"))

# Set the factor levels of lda.preds to match bc$diagnosis
complex.preds <- factor(complex.preds, levels = levels(training$fracture))

confusion_complex <- confusionMatrix(data = complex.preds, reference = training$fracture, positive = "Yes")

confusion_complex
```

```{r}
#library(dummies)
#pc.result2<-prcomp(bone[, -(1:3)],scale.=T)
bone_exclude <- c(1:3, which(names(bone)=="fracture"))
bone_select <- bone[, -bone_exclude]
bone_encoded <- data.frame(sapply(bone_select, function(col) if(is.factor(col)) as.numeric(col) else col))


pc.bone.result<-prcomp(bone_encoded, scale.=T)
pc_bone_data <- data.frame(pc.bone.result$x, fracture = bone$fracture)
#head(pc_bone_data)

#scree plot for optimization
ideal_variance <- 0.9
eigenvals <- pc.bone.result$sdev^2
plot(eigenvals/sum(eigenvals), type="l", main="Scree Plot", ylab="Prop. Var. Explained", xlab="PC #", ylim=c(0,1))
cumulative.prop <- cumsum(eigenvals/sum(eigenvals))
lines(cumulative.prop, lty=2)
abline(h=ideal_variance, col="red", lty=2)
pc_number <- which(cumulative.prop >= ideal_variance)[1]
message <- paste0("For ", ideal_variance*100, "% of total variance, ", pc_number, " Principle Components required.\n\n", sep="")

abline(v=pc_number, col="blue", lty=3)

# Add the message to the x-axis label
mtext(message, side=1, line=4)

```

```{r}

pc_select <- data.frame(pc.bone.result$x[, 1:8], fracture = pc_bone_data$fracture)
#head(pc_select)


complex_model_pca<-train(fracture~.
             ,
             data=pc_select,
             method="lda",
             trControl=fitControl,
             metric="logLoss")

predictions <- predict(complex_model_pca, pc_select, type = "prob")[,"Yes"]

#Getting confusion matrix
threshold = 0.5
complex.pca.preds <- factor(ifelse(predictions > threshold, "No", "Yes"),
                    levels = c("No", "Yes"))

complex.pca.preds <- factor(complex.pca.preds, levels = levels(pc_select$fracture))

head(complex.pca.preds)

confusionMatrix(data = complex.pca.preds, reference = pc_select$fracture, positive = "Yes")
```

