---
title: "Project 2 EDA"
author: 'MSDS 6372: Rob Lane'
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

```{r}
library(aplore3)
#head(glow_bonemed)
```

glow_bonemed variables:

| **Variable** | **Type**     | **Description**                                                                                                                                 |
|--------------|--------------|-------------------------------------------------------------------------------------------------------------------------------------------------|
| bonemed      | 2-factor     | Bone medications at enrollment (1: No, 2: Yes)                                                                                                  |
| bonemed_fu   | 2-factor     | Bone medications at follow-up (1: No, 2: Yes)                                                                                                   |
| bonetreat    | 2-factor     | Bone medications both at enrollment and follow-up (1: No, 2: Yes)                                                                               |
| sub_id       | int          | Identification Code (1 - n)                                                                                                                     |
| site_id      | int          | Study Site (1 - 6)                                                                                                                              |
| phy_id       | int          | Physician ID code (128 unique codes)                                                                                                            |
| priorfrac    | 2-factor     | History of Prior Fracture (1: No, 2: Yes)                                                                                                       |
| age          | int          | Age at Enrollment (Years)                                                                                                                       |
| weight       | num          | Weight at enrollment (Kilograms)                                                                                                                |
| height       | int          | Height at enrollment (Centimeters)                                                                                                              |
| bmi          | num          | Body Mass Index ($\frac{kg}{m^2}$)                                                                                                              |
| premeno      | 2-factor     | Menopause before age 45 (1: No, 2: Yes)                                                                                                         |
| momfrac      | 2-factor     | Mother had hip fracture (1: No, 2: Yes)                                                                                                         |
| armassist    | 2-factor     | Arms are needed to stand from a chair (1: No, 2: Yes)                                                                                           |
| smoke        | 2-factor     | Former or current smoker (1: No, 2: Yes)                                                                                                        |
| raterisk     | 3-factor     | Self-reported risk of fracture (1: Less than others of the same age, 2: Same as others of the same age, 3: Greater than others of the same age) |
| fracscore    | int          | Fracture Risk Score (Composite Risk Score)                                                                                                      |
| **fracture** | **2-factor** | **Any fracture in first year (1: No, 2: Yes)**                                                                                                  |

```{r, echo=F, fig.width=7.5, fig.height=7.5}
library(graphics)
bone <- glow_bonemed
#Survived = 1, Perished = 0.  Therefore, salmon means perished, turquoise means survived.
class_colors <- c("salmon", "turquoise")

#Excluding IDs because they don't describe the problem.
exclude_vars <- c("sub_id", "site_id", "phy_id")
include_cols <- setdiff(names(bone), exclude_vars)

pairs(bone[, include_cols],
      col = class_colors[bone$fracture])
```

**Multicollinearity Found:**  
fracscore : age  
bmi : weight  
site_id : phy_id  *


